---
title: "607 Homework 1"
author: "Armenoush Aslanian-Persico"
date: "February 7, 2016"
output: html_document
---

##Load libraries.

```{r load-data, eval=TRUE}
library(devtools)
library(RCurl)
library(plyr)
```

##Import file as CSV, check variables.
Data has no header row so we must indicate "header=F".
```{r}
x <- getURL("https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version1")
y<-data.frame(read.csv(text=x, header=F))
dim(y)
head(y)
```

##Rename column headers.
```{r}
head(rename(y,c("V1"="ID","V2"="River","V3"="Location","V4"="Erected","V5"="Purpose","V6"="Length", "V7"="Lanes", "V8"="Clear-G","V9"="T or D","V10"="Material","V11"="Span","V12"="Rel-L","V13"="Type")))
```

##Select a subset by row.
###How many wood bridges are over 100 years old?
```{r}
OldWoodenBridges<- subset(y, V4 <1915 & V10=="WOOD")
OldWoodenBridges
summary(OldWoodenBridges)
```
This returns 16 bridges.  

###How many steel bridges are over 100 years old?
```{r}
OldSteelBridges<- subset(y, V4 <1915 & V10=="STEEL")
OldSteelBridges
summary(OldSteelBridges)
```
This returns 38 bridges.  

##Select a subset by column and row.
###Show select information for new bridges.
```{r}
NewBridges<-subset(y,V4>1960,select=c(V1,V4,V9,V10,V11,V13))
NewBridges
summary(NewBridges)
```

##Bar plots
###What are the most common bridge materials in Pittsburgh?
```{r}
barplot(table(y$V10), ylim=c(0,108),main="Pittsburgh Bridges by Material")
```
This makes sense considering Pittsburgh's steel history.  

###How many bridges were built by year?
```{r}
barplot(table(y$V4), ylim=c(0,10),main="Pittsburgh Bridges by Year")
```

