---
title: "Part 1: Merge datasets; export complete cases only."
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r 0-libraries, warning=FALSE, message=FALSE}
library(proto)
library(gsubfn)
library(RSQLite)
library(knitr)
library(data.table)
library(plyr)
library(dplyr)
library(sqldf)


#abbreviations
#dfp is the civil list/payroll dataset
#dft is the titles dataset
```


# Unique Titles

``````{r 1-uniquetitles}
#this imports a list of unique titles previously created from the titles dataset

dft <- data.frame(read.csv("20180305titles.csv"),stringsAsFactors = FALSE)

dft %>%
    arrange((UniqueTitle))%>%
  head(5)%>%
  kable
str(dft) #4424 obs. of 10 variables
```

# How many titles in the Unique Titles dataset?
> 4424 unique titles

# Employee List (Civil List)

``````{r 2-civillist}
#import the payroll dataset

dfp <- data.frame(read.csv("20180305payroll.csv"), stringsAsFactors = FALSE)
str(dfp)

dfp%>%
  head(5)%>%
  kable

#count of employees by year in the payroll dataset
byyear <- group_by(dfp,CALENDAR.YEAR)
( sumyear <- summarize(byyear,count=n()) )

#commented out
#count of employees by agency in the payroll dataset
#byagency <- group_by(dfp,ADDRESS)
#( sumyear <- summarize(byagency,count=n()) )


#count of employees by agency in the payroll dataset
#different method
#sort by total number of employees in years 2014+2015
dfp %>%
     group_by(ADDRESS) %>%
     summarise(total.count=n()) %>%
     arrange(desc(total.count))

#str(dfp) #570941 obs. of 7 variables
```

# Join Two Tables

``````{r 3-join}

#rename columns in both datasets
setnames(dft, "UniqueTitle", "TITLE")
setnames(dfp, "TTL..", "TITLE")
setnames(dfp, "SAL.RATE", "SAL")
setnames(dfp, "CALENDAR.YEAR", "YEAR")

#left outer join to match by title
dfm1<-merge(x = dfp, y = dft, by = "TITLE", all.x = TRUE)
#dfm1<-data.frame(dfm1, stringsAsFactors = FALSE)


dfm1 %>% mutate_if(is.factor, as.character) -> dfm1
str(dfm1)

#show most frequent descriptions in descending order; 153485 rows did not have a matching title code
dfm1 %>%
     group_by(Description) %>%
     summarise(total.count=n()) %>%
     arrange(desc(total.count))

```

# What rows are we excluding?
## Incomplete cases: Employees listed on the payroll dataset but whose job is not listed on the titles dataset

``````{r 4-employees-with-no-title-match}
dfna <- dfm1[is.na(dfm1$Description),] #should be 153485
#dfna<-data.frame(dfna, stringsAsFactors = FALSE)
str(dfna)
#first attempt, with factors
#153485 rows of 570491 are unmatched
#13598 mayoral
#second attempt, stringsasfactors-false

# what agencies are we excluding? Note the counts here are the count of employee payroll years. An employee on payroll in 2014 and 2015 would be summarized here as 2 separate observations.

dfna %>%
     group_by(ADDRESS) %>%
     summarise(total.count=n()) %>%
     arrange(desc(total.count))%>%
  head(70)%>%
  kable

# how many records per year are we excluding?
dfna %>%
     group_by(YEAR) %>%
     summarise(total.count=n()) %>%
  head(5)%>%
  kable

#write.csv(dfna, file = "dfna.csv")
```

# View joined table; complete cases only

```{r 5-join}

#show complete cases only-- do not include NA rows
dfc<-dfm1[complete.cases(dfm1), ]

#convert salary to numeric, remove currency symbol
dfc$SAL <- as.numeric(gsub('[$,]', '', dfc$SAL))

#show head by descending salary
dfc %>%
    arrange(desc(SAL))%>%
  head(5)%>%
  kable

# check count of rows in dfc. should be total num minus na. Should be 570491-153485 = 417006
str(dfc)
```

# Export complete cases
```{r 5a}
write.csv(dfc, file = "dfc.csv")
```


---
# The following sections will be in Part 2 and are listed here for reference purposes only:

# Show unique values in Pay Type Column

```{r 6, eval=FALSE}
#bypaytype <- group_by(dfc,PayType)
#( sumpaytype <- summarize(bypaytype,count=n()) )

bypaytype<- dfc %>% count(PayType, YEAR) %>% ungroup() %>% arrange(desc(n))

kable(head(bypaytype,20))
```

# Show unique values in SalaryMinMaxValid Column

```{r 7, eval=FALSE}
#byvalidsal <- group_by(dfc,SalaryMinMaxValid)
#( sumvalidsal <- summarize(byvalidsal,count=n()) )

byvalidsal<- dfc %>% count(SalaryMinMaxValid, YEAR) %>% ungroup() %>% arrange(desc(n))

kable(head(byvalidsal,20))

```

# Show unique values in TitleStatus Column

```{r 8, eval=FALSE}
#bytitlestatus <- group_by(dfc,TitleStatus)
#( sumtitlestatus <- summarize(bytitlestatus,count=n()) )

bytitlestatus<- dfc %>% count(TitleStatus, YEAR) %>% ungroup() %>% arrange(desc(n))

kable(head(bytitlestatus,20))

```

# Show unique values in UnionName Column

```{r 9, eval=FALSE}
#byunion <- group_by(dfc,UnionName)
#( sumunion <- summarize(byunion,count=n()) )

byunion<- dfc %>% count(UnionName, YEAR) %>% ungroup() %>% arrange(desc(n))

kable(head(byunion,20))

```


# Show duplicated names

```{r 11, eval=FALSE}
#kable(duplicated(dfc$NAME))
```
# Investigate the EitherAnnualHourly rows to determine salary category.

# Narrow the dataset active titles only, valid salary only, annual salary only.

# Issues with data integrity

> Describe creation of indicator columns: CountSubTitles, Min.MinRate, Max.MaxRate, SalaryMinMaxVlaid, MultipleSubtitles, PayType, PayPeriod

> Typo in at least one record, showing $350,000 salary minimum which should be $35,000


> Salary rate is not always within the min/max allowable of the title. This may be because the salary may include overtime, or the title list may have outdated salary amounts.

> 153485 rows of 570491 are unmatched, meaning there are 153485 employee records where the title code does not correspond to a published title in the civil list. These rows will have to be removed from the dataset because no title analysis will be possible.