---
title: 'Part 3: Visualizations'
output:
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries

```{r 0-libraries, warning=FALSE, message=FALSE}
library(proto)
library(gsubfn)
library(RSQLite)
library(knitr)
library(data.table)
library(plyr)
library(dplyr)
library(sqldf)
library(ggplot2)
library(RColorBrewer)
library(grid)
library(ggrepel)
library(plotly)
library(kableExtra)
```


# Donut chart:  Top 10 agencies with the most jobs under $10K, 2015

```{r 1}
df1<- data.frame(read.csv("agency15low.csv"),stringsAsFactors = FALSE)

df1 %>% mutate_if(is.factor, as.character) -> df1

p1 <- df1 %>%
  plot_ly(labels = ~ADDRESS, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Agencies with the most jobs under $10K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p1

```


# Donut chart:  Top 10 titles with the most jobs under $10K, 2015

```{r 2}
df2<- data.frame(read.csv("agency15low-titles.csv"),stringsAsFactors = FALSE)

df2 %>% mutate_if(is.factor, as.character) -> df2

p2 <- df2 %>%
  plot_ly(labels = ~Description, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Titles with the most jobs under $10K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p2

```


# Donut chart:  Top 10 agencies with the most jobs over $10K, 2015

```{r 3}
df3<- data.frame(read.csv("agency15high.csv"),stringsAsFactors = FALSE)

df3 %>% mutate_if(is.factor, as.character) -> df3

p3 <- df3 %>%
  plot_ly(labels = ~ADDRESS, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Agencies with the most jobs over $10K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p3

```


# Donut chart: Top 10 titles with the most jobs over $10K, 2015

```{r 4}
df4<- data.frame(read.csv("agency15high-titles.csv"),stringsAsFactors = FALSE)

df4 %>% mutate_if(is.factor, as.character) -> df4

p4 <- df4 %>%
  plot_ly(labels = ~Description, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Titles with the most jobs over $10K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p4

```

# Histogram: Distribution of 10K+ salaries

```{r 5}
df15<- data.frame(read.csv("df15.csv"),stringsAsFactors = FALSE)

hist(df15$SAL, 
     main ="Distribution of salaries over $10K, 2015",
     xlab="Salary",
     border="black",
     col="#58D68D",
     xlim=c(0,200000),
     ylim=c(0,50000),
     breaks = seq(0, 350000, by = 10000)
     #prob=TRUE
     #breaks=c(0,10000,20000,30000,40000,50000,60000,70000,80000,90000,100000,110000,120000,130000,350000)
     )

#lines(density(df15$SAL))


```

# Scatterplot: Count of jobs in title vs. Average salary per title

```{r 6}
sal15a<-df15 %>%
  group_by(Description) %>%
  mutate(Count = n()) %>%
  group_by(TITLE, Description, Count) %>%
  summarise_at(vars(SAL), funs(round(mean(., na.rm=TRUE),0)  ))%>%
  arrange(desc(Count))

df5<-head(sal15a,100)

p5 <- plot_ly(
  data = df5, 
  x = ~Count, 
  y = ~SAL, 
  type = 'scatter', 
  mode = 'markers',
  hoverinfo='text',
  text = ~paste(
    'Title: ', Description,
    '<br> Employees: ', Count,
    '<br> Avg Salary: ', SAL)) %>%
  layout(title="100 most popular titles with average salary by title")
p5


p5 <- plot_ly(
  data = df5, 
  x = ~Count, 
  y = ~SAL, 
  type = 'scatter', 
  mode = 'markers',
  hoverinfo='text',
  text = ~paste(
    'Title: ', Description,
    '<br> Employees: ', Count,
    '<br> Avg Salary: ', SAL)) %>%
  layout(
    xaxis = list(range = c(0, 5000)),
    yaxis = list(range = c(0, 170000)),
    title="Popular titles with average salary by title, <5000 serving in title")
p5

```

# Table: Do some agencies have a greater variety of job titles available?

The following is a count of unique titles by agency.

```{r 7}
df6<-df15

df6 %>%
  group_by(ADDRESS) %>%
  summarise(count = n_distinct(TITLE))%>% 
  arrange(desc(count))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 
```

# Table: Which titles have employees across the most agencies?

The high counts here are skewed due to each Community Board being represented as a different city agency.

```{r 8}
df7<-df15

df7 %>%
  group_by(Description) %>%
  summarise(count = n_distinct(ADDRESS))%>% 
  arrange(desc(count))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 


```

# Table: Which titles have employees across the most agencies, excluding agencies which are Community Boards?

These could be considered the most flexible titles, for employees who may want to move around across agencies while retaining their civil service title.

```{r 8a}

df7a<-df7[- grep("COMMUNITY BOARD", df7$ADDRESS),]

df7a %>%
  group_by(Description) %>%
  summarise(count = n_distinct(ADDRESS))%>% 
  arrange(desc(count))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 
```

# Bar Plot: Which titles have employees across the most agencies, excluding agencies which are Community Boards?

```{r 8b, eval=FALSE}
# This keeps all 1089 descriptions from the original data frame
df7b<-df7a %>%
  group_by(Description) %>%
  summarise(count = n_distinct(ADDRESS))%>% 
  arrange(desc(count))%>%
  top_n(10)%>%
  as.data.frame()

str(df7b)

p7 <- plot_ly(
  data = df7b, 
  x = ~Description, 
  y = ~count, 
  type = 'bar', 
  #mode = 'markers',
  hoverinfo='text',
  text = ~paste(
    'Title: ', Description,
    '<br> Employees: ', count) ) %>%
  layout(title="Titles across the most agencies, excluding Community Boards")

p7
```

```{r 8c}

df7b<- data.frame(read.csv("titlesmultipleagencies.csv"),stringsAsFactors = FALSE)

df7b$Description <- factor(df7b$Description, levels = unique(df7b$Description)[order(df7b$COUNT, decreasing = TRUE)])


p7 <- plot_ly(
  data = df7b, 
  x = ~Description, 
  y = ~COUNT, 
  type = 'bar', 
  hoverinfo='text',
  color="#D7BDE2",
  text = ~paste(
    'Title: ', Description,
    '<br> Agencies: ', COUNT) ) %>%
  layout(title="Titles across the most agencies, excluding Community Boards", margin = list(b = 200), xaxis = list(tickangle = 90))

p7

```

# For titles across multiple agencies, do some agencies hire the same title at a higher average salary?

# What titles have the most opportunities for salary increases?
## Min and max in each title

# What are the most significant changes between 2014 and 2015?

# Is there a correlation between union representation and salary?

# Can this data predict current job postings in 2018? 

# Is there a relationship between an agency's budget and its average salaries?